---
title: "Version Control and Collaboration with Git and GitHub"
subtitle: "Resul Umit"
institute: ""
date: "March 2021"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
      ratio: "16:9"
    seal: false
---

class: inverse, center, middle

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, crayon.enabled = TRUE)

library(crayon)
library(fansi)
library(dplyr)
library(ggplot2)
library(stargazer)
library(countdown)
```

<style type="text/css">

.hljs-github .hljs {
    background: #e5e5e5;
}

.inline-c, remark-inline-code {
   background: #e5e5e5;
   border-radius: 3px;
   padding: 4px;
   font-family: 'Source Code Pro', 'Lucida Console', Monaco, monospace;
}


.yellow-h{
   background: #ffff88;
}


.out-t, remark-inline-code {
   background: #9fff9f;
   border-radius: 3px;
   padding: 4px;
   
}

.pull-left-c {
    float: left;
    width: 58%;
    
}

.pull-right-c {
    float: right;
    width: 38%;
    
}

.medium {
    font-size: 75%
    
}

.small {
    font-size: 50%
    }

.action {
    background-color: #f2eecb;
  
}


</style>


# Version Control and Collaboration with Git and GitHub

<br>

### Resul Umit

### March 2021

.footnote[

[Skip intro &mdash; To the contents slide](#contents-slide).

]

---
## Who am I?

Resul Umit

- post-doctoral researcher at the University of Oslo

- interested in representation, elections, and parliaments
    - [a recent publication](https://doi.org/10.1177%2F1478929920967588):
    Parliamentary communication allowances do not increase electoral turnout or incumbents’ vote share

--

<br>

- teaching workshops also on

  - [writing reproducible research papers with R](https://resulumit.com/blog/rmd-workshop/)
  - [working with Twitter data in R](https://resulumit.com/syllabi/21spring_twtr.pdf)
  - [creating professional websites with R](https://resulumit.com/syllabi/20spring_rbd.pdf)
    
--

<br>

- more information available at [resulumit.com](https://resulumit.com/)

---
## The Workshop &mdash; Overview

- Half a day, on how to 

  - version our academic work
  - store the different versions securely &mdash; in multiple, separate places
  - collaborate with co-authors &mdash; in writing as well as versioning and storing

--

<br>

- Using Git and GitHub, through
   
  - command line tools
  - GitHub Desktop
  - R and RStudio
  
--

<br>

- Most useful for projects based on pain text
   - e.g., Markdown, rather than Microsoft Word
   - but not exclusively so 
       - projects based on binary files would also benefit
       - you may switch to plain text in the future
        
---
## The Workshop &mdash; Motivation

- Research projects have many versions, as they are typically completed

  - over a long period of time
  - in numerous sittings
  - by multiple authors 
  - on multiple computers

--

<br>

- With many versions created over time, in different sittings ..., there emerges various challenges

   - keeping track of changes and versions
   - reverting to a previous version when necessary
   - storing versions safely in multiple locations
   - communicating the versions to other authors and/or computers
   - working on the same project with co-authors at the same time

--

<br>

- We all have to deal with these challenges on daily basis
   
   - efficient workflows can save precious resources
      - e.g., time, but also money and work itself from being wasted
   
---
## The Workshop &mdash; Worklow

- Using Git and GitHub 
   - originally designed for software development, but can be adopted to academic work
       - Git creates, saves, and tracks versions
           - can also save them elsewhere, such as on GitHub
       - Github keeps a copy online
           - allowing for safe storage, collaboration
  - is daunting to beginners, but offers many benefits over alternatives
--

<br>

- Alternatives are easier for beginners, but sub-optimal in the long run

   - e.g., editing, renaming, saving, and emailing
   - e.g., using applications or websites such as Dropbox, Google Docs, Overleaf

---




















---
## Version Control &mdash; Git and GitHub &mdash; Definitions

- Git
    - a software that keeps track of versions of a set of files
    - it is *local* to you, the records are kept on your computer

--

- GitHub
    - a hosting service, or a website, that can keep the records
    - it is *remote* to you, like the Dropbox website
    - but unlike Dropbox, GitHub is specifically structured to keep records with Git

--

- Repository, or repo
    - a set of files whose records are kept together, by Git and/or on GitHub
    - it is like a folder, which can keep files and other folders containing files

---
## Version Control &mdash; Git and GitHub &mdash; Definitions

- To commit
    - to take a snaphot of, or to version, a repository
    - it is like saving a new version of all files and sub-folders in your project folder with a new name
    - it is local, the records are kept on your computer unless you `push`

--

- To push
    - to move the records from Git to GitHub, from your computer to online server
    - it is like uploading (the new versions of) your files and sub-folders to a website, 
    - it also involves merging, if this not the first push

.footnote[

<sup>*</sup> For projects that are single-authored on a single computer, merging is typically automatic. It becomes an issue for collaborated projects, which we will cover in the next section &mdash; Part 9.

]


## Collaboration

- Many research papers are written by multiple authors and/or on multiple computers

    - yourself on a different computer (e.g., laptop at home, desktop at office), poses similar challenges as collaboration
    
--

- With multiple authors and/or computers, there emerges at least two additional challenges beyond version control

   - communicating the versions to other authors and/or computers
   - working on the same project with co-authors at the same time

--

- We all manage collaboration, in different ways, such as

   - edit, rename, save, e-mail
   - use applications or websites such as Dropbox, Google Docs, Overleaf
   - use distributed version control systems such as Git and GitHub
   
---
## Collaboration &mdash; Git and GitHub &mdash; Definitions

- To pull

    - to move the (presumably) up-to-date records from GitHub to your computer
    - it is like downloading a zipped folder of files

--

- To merge

    - to integrate different versions into a single version
        - e.g., the old version on your laptop, with (the changes in) the new version from GitHub
    - except the first push or pull, pushing and pulling necessiate merging
    
--

- Merge conflict

    - emerges when versions to be merged include edits *on the same line of the same file*
        - edits on different lines are not a problem as changes are tracked line by line    
<br>
    - less likely to occur in one-author-multiple-computer setting 
        - more likely while collaborating with others    
<br>
    - requires human intervention, to decide which edit to keep and which one to discharge

---
## Collaboration &mdash; Git and GitHub &mdash; Definitions

- Branch

    - a line of development in a repository; a copy of the repository, with all its versions, at a given time
    - by default, repositories have one branch, called *master*


- Pull request    

    - a proposal to pull and merge
        - e.g., a proposal from one co-author to another, 
        -e.g., tp merge a branch into master
    - it allows a review of changes on GitHub before merge, to deal with potential merge conflicts
    





   




---
## The Workshop &mdash; Overview

- Two days, on how to write reproducible research papers with R Markdown

  - 200+ slides, 40+ exercises, and time for converting a real project

--

- Based on converting a mock manuscript written in Word to R Markdown

  - plus, improving its reproducibility and version-controlling it
  - with a PDF output in mind

--

- Designed for researchers with basic knowledge of R programming language

  - does not cover programming with R
      - e.g., writing functions    
<br>
  - ability to regress, plot, and table in R will be very helpful
      - but not absolutely necessary &mdash; these skills can be developed after learning R Markdown as well
        
---
name: contents-slide

## The Workshop &mdash; Contents

[Part 1. Getting the Tools Ready](#part1)
   - e.g., downloading course material

[Part 2. Command Line](#part2)
   - e.g., creating a new document

[Part 3. GitHub Desktop](#part3)
   - e.g., defining output format

[Part 4. R and RStudio](#part4)
   - e.g., adding emphasis to text
   

---
## The Workshop &mdash; Organisation

- Sit in groups of two

  - participants learn as much from their partner as from instructors
  - play the role of co-authors
  
--

- Type, rather than copy and paste, the code that you will find on these slides

  - typing is a part of the learning process
  
--

- When you have a question

  - ask your partner
  - google together
  - ask me

---
class: action

## The Workshop &mdash; Organisation &mdash; Slides

Slides with this background colour indicate that your action is required, for

- setting the workshop up
    - e.g., downloading course material 
    
- completing the exercises
    - e.g., saving a different version of a file
    - there are 20+ exercises
    - these slides have countdown timers
    
    
`r countdown(minutes = 3, seconds = 00, top = 0)`

---
## The Workshop &mdash; Organisation &mdash; Slides

- Codes and texts that go in .inline-c[appear as such &mdash; in a different font, on gray background]
    - long codes and texts will have their own line(s)

```md

ggplot(data = df, mapping = aes(x = h5_median, y = h5_index, color = subfield)) +
       geom_point() +
       facet_wrap(. ~ branch) +
       scale_colour_discrete(name = "Journal Type", breaks = c(0, 1), labels = c("Generalist", "Subfield"))    

```

---
## The Workshop &mdash; Organisation &mdash; Slides

- Codes and texts that go in .inline-c[appear as such &mdash; in a different font, on gray background]
    - long codes and texts will have their own line(s)

- Results that come out in .out-t[appear as such &mdash; in the same font, on green background]
    - except very obvious results, such as screenshots
    
--
    
- Specific sections are .yellow-h[highlighted yellow as such] for emphasis
    - these could be for anything &mdash; codes and texts in input, results in output, and/or texts on slides
    
--

- The slides are designed for self-study as much as for the workshop
    - *accessible*, in substance and form, to go through on your own

---
## The Workshop &mdash; Aims

- To make you aware what is possible with Git and GitHub

  - we will cover a large breath of issues, not all of it is for long-term memory
      - one reason why the slides are designed for self study as well    
<br>
  - awareness of what is possible, `Google`, and perseverance are all we need

--

- To encourage you to convert into the Git and GitHub workflow 

  - practice with a mock project, co-author
  - start converting a real one right after the workshop
      - happy to help via emails afterwards
    
---
name: part1
class: inverse, center, middle

# Part 1. Getting the Tools Ready

.footnote[

[Back to the contents slide](#contents-slide).

]

---
## Workshop Materials &mdash; Overview

Materials, available at <https://github.com/resulumit/git_workshop/tree/materials>, have the following structure

```

git_workshop-materials
   |
   |- data
   |  |
   |  |- journals.csv
   |
   |- manuscript
   |  |
   |  |- journals.docx
   |  |- journals.pdf
   |  |- journals.Rmd
   |
   |- README.md

```

---
name: download-zip
class: action

## Course Materials &mdash; Download from the Internet

- Download the materials from <https://github.com/resulumit/git_workshop/tree/materials>

  - on the webpage, follow

> `Code -> Download ZIP`

<br>

- Unzip and rename the folder

  - unzip to a location that is not synced
     - e.g., perhaps Documents, but not Dropbox
  
  - rename the folder as `YOURNAME_git_workshop`
     - e.g., `resul_git_workshop`
     - this renaming will come handy later on
  
---
## Course Materials &mdash; Contents

- `manuscript\paper.docx`

    - the document, formatted in Word but saved as PDF, that we will re-create with R Markdown 
    - randomly generated sentences, with figures and tables from randomly a generated dataset<sup>*</sup>
    - key sections in-need of attention are highlighted yellow
    
---
## Course Materials &mdash; Contents  

`data\journals.csv`
- a dataset created with the `fabricatr` package (Blair et al., 2019), imagined to explore the `Google Scholar` rankings of fictitious journals
    
- includes the following variables 
    - **name**: journals (1090 random titles)
    - **origin**: geographic origins (five continents)
    - **branch**: major discipline of journals (four branches)
    - **since**: time of first publication (years)
    - **h5_index**: H5 Index (integers)
    - **h5_median**: H5 Median (integers)
    - **english**: English (1) *vs.* other-language (0) journals
    - **subfield**: subfield (1) *vs.* generalist (0) journals
    - **issues**: number of issues published per year (integers)

---
## Course Materials &mdash; Contents  

- `image\google_scholar.png`

    - a screeenshot image of the [Google Scholar homapage](https://scholar.google.ch/)

---
class: action
name: install-git

## Git &mdash; Download from the Internet and Install

- For Windows, install 'Git for Windows', downloading from [https://gitforwindows.org](https://gitforwindows.org/)
       - select 'Git from the command line and also from 3rd-party software'
       
- For Mac, install 'Git', downloading from [https://git-scm.com/downloads](https://git-scm.com/downloads)

---
class: action

## GitHub &mdash; Open an Account

Sign up for GitHub at [https://github.com](https://github.com/)

- registering an account is free

- usernames are public
      - either choose an anonymous username (e.g., `asdf029348`)
      - or choose one carefully &mdash; it becomes a part of users' online presence
      
- usernames can be changed later
        
---
## Other Resources<sup>*</sup>

- Pandoc User's Guide

   - available at [https://pandoc.org/MANUAL.html](https://pandoc.org/MANUAL.html)

- R Markdown: The Definitive Guide (Xie et al., 2019)

   - open access at [https://bookdown.org/yihui/rmarkdown](https://bookdown.org/yihui/rmarkdown/)
    

- R for Data Science (Wickham and Grolemund, 2019)
  
   - open access at [https://r4ds.had.co.nz](https://r4ds.had.co.nz/)
   
.footnote[

<sup>*</sup> During the workshop, R Markdown Cheat Sheet is likely to be more helpful than these resources, which I recommend to be consulted after the workshop.

]



---
name: part8
class: inverse, center, middle

# Part 8. Using Version Control

.footnote[

[Back to the contents slide](#contents-slide).

]

---


---
## Version Control &mdash; Git and GitHub

Version control with Git and GitHub requires

1. .yellow-h[initial setup], done once<sup>*</sup>
    - unless for a new computer or, if ever, a new GitHub account
    - a bit technical, but worth the hassle
    
2. .yellow-h[project setup], repeated for every paper
    - shorter, less complicated

.footnote[

<sup>*</sup> We have started this process already, in Part 1 of the workshop, by downloading and installing Git and signing up for GitHub. [Back to the relevant slide](#install-git).

]

---
class: action

## Version Control &mdash; Git &mdash; Inital Setup

1) Enable version control with RStudio

- from the RStudio menu, follow:

> `Tools -> Global Options -> Git/SNV -> Enable version control interface for RStudio projects`

- RStudio will likely find Git automatically. In case it cannot, Git is likely to be at

    - `c:/Program Files/Git/bin/git.exe` on Windows

    - `/usr/local/git/bin/git` on Mac

---
class: action

## Version Control &mdash; Git &mdash; Inital Setup


2) Set Git Bash as your shell (Windows-only step)

- from the RStudio menu, follow:

> `Tools -> Global Options -> Terminal -> New terminals open with: Git Bash`

---
class: action

## Version Control &mdash; Git &mdash; Inital Setup

3) Introduce yourself to Git

- from the RStudio menu, follow:

> `Tools -> Terminal -> New Terminal`

- enter the following lines in the Terminal, with the email address that you have used to sign up for GitHub

```yaml
git config --global user.name "YOUR-NAME" 
git config --global user.email "YOUR-EMAIL-ADDRESS"
```

- enter the following line in the Terminal, to observe whether the previous step was sucessful

```yaml
git config --global --list

```

---
class: action

## Version Control &mdash; Git and Github &mdash; Project Setup<sup>*</sup>

1) Initiate local version control with Git

- from the RStudio menu, follow:

> `Tools -> Version Control -> Project Setup... -> Version Control System -> Git`


- after confirming your new repository, and restarting the session, observe that 

   - now there is now a Git tab in RStudio, documenting the differences between you local repository and the one on GitHub. When you change a file, it will appear here.  
   
   - your project now includes a `.gitignore` file
       - this is where you can list files and/or folders to be excluded from being tracked

.footnote[

<sup>*</sup> These instructions presume there is an exiting RStudio project to be set up for version control. If not, or to start a new project, follow from [this slide](#rstudio-project) first.

]

---
class: action

## Version Control &mdash; Git and Github &mdash; Project Setup

2) Create a new GitHub repository

- on GitHub, follow:

> `Repositories -> New -> Repository name (e.g., "rwd_workshop") -> Public -> Create repository`

- observe that

    - repository URLs have the following structure: https://github.com/USER_NAME/REPOSITORY_NAME
    
        - this is the address to view the repository online
        - for use in the `Terminal`, the address gets the `.git` extension
            - e.g., https://github.com/USER_NAME/REPOSITORY_NAME.git

---
class: action

## Version Control &mdash; Git and Github &mdash; Project Setup

3) Push an existing repository

- from the RStudio menu, follow:

> `Tools -> Terminal -> New Terminal`

- enter the following lines in the `Terminal`, with your username and repository name

```yaml
git remote add origin https://github.com/USER_NAME/REPOSITORY_NAME.git
git add .
git commit -m "first commit"
git push -u origin master

```

--

- if this is your first time using GitHub with RStudio, you will be prompted to authenticate
   - follow the instructions on your screen and in your email

- observe that your project files are now online, listed on the GitHub repository

---
class: action
name: version-control-workflow

## Version Control &mdash; Git and Github &mdash; Workflow

1) Edit and Save 

- work on one or more files under version control
    - e.g., delete the first sentence of the abstract in `journals.Rmd`, and save it
    - under the Git tab in RStudio, find the list of files that you edited since the last push
    - these will have `M`, for modified, as `Status`     

--

2) Commit and Push

- tick `Staged`<sup>*</sup>  for one or more files that you would like to commit
   - enter a `Commit message` that summarises the edits
   - click `Commit` to create a record of the new version locally to your computer
   - click `Close -> Push` to push the version to GitHub

.footnote[

<sup>*</sup> To stage is to add files to be comitted. It allows us to commit files individually or together with other files.

]

---
class: action
name: git-github-basic-workflow

## Version Control &mdash; Git and Github &mdash; Workflow

1) Edit and Save 

- work on one or more files under version control
    - e.g., delete the first sentence of the abstract in `journals.Rmd`, and save it
    - under the Git tab in RStudio, find the list of files that you edited since the last push
    - these will have `M`, for modified, as `Status`     

2) Commit and Push

- tick `Staged` for one or more files that you would like to commit
    - enter a `Commit message` that summarises the edits
    - click `Commit` to create a record of the new version locally to your computer
    - click `Close -> Push` to push the version to GitHub

- observe the changes in the Git tab in RStudio and on the GitHub repository

---
## Version Control &mdash; Git and Github &mdash; `.gitignore`

- `.gitignore` specifies which file(s) and/or folder(s) should be excluded from version control

    - a set of project-specific files are ignored by default
        - see your `.gitignore` file

--

- `.gitignore` lists one item per line

    - each line has a pattern, which determines whether one or more files or folders are to be ignored

--

- See the documentation at <https://git-scm.com/docs/gitignore> 

    - for pattern formats and further details

---
## Version Control &mdash; Git and Github &mdash; `.gitignore`

- There might be good reasons to ignore some others, including files

    - that contain information that we do not want others to see
        - e.g., personal API keys    
<br>
    - that we do not have the right to share with others
        - e.g., secondary data with user agreements otherwise    
<br>
    - that we (re-)create automatically as outputs
        - e.g., `journals.pdf`, as opposed to journals.Rmd
        
---
## Version Control &mdash; Git and Github &mdash; `.gitignore`

.pull-left[

- Observe that, by default, `.gitignore` has a list of project-specific files
    - you can delete, or comment out, any or all to start including them in version control

]

.pull-right[

```{r, eval = FALSE}
.Rproj.user
.Rhistory
.RData
.Ruserdata
```

]

---
## Version Control &mdash; Git and Github &mdash; `.gitignore`

.pull-left[

- Observe that, by default, `.gitignore` has a list of project-specific files

- In addition, you can ignore, for example,

    - <span style="background-color: #ffff88;">a specific folder</span>, relative to the root directory

]

.pull-right[

```{r, eval = FALSE}
.Rproj.user
.Rhistory
.RData
.Ruserdata
/manuscript/      #<<
```

]

---
## Version Control &mdash; Git and Github &mdash; `.gitignore`

.pull-left[

- Observe that, by default, `.gitignore` has a list of project-specific files

- In addition, you can ignore, for example,

    - a specific folder, relative to the root directory

    - <span style="background-color: #ffff88;">a specific file in a specific folder</span>, relative to the root directory

]

.pull-right[

```{r, eval = FALSE}
.Rproj.user
.Rhistory
.RData
.Ruserdata
/manuscript/
/manuscript/journals.pdf         #<<
```

]

---
## Version Control &mdash; Git and Github &mdash; `.gitignore`

.pull-left[

- Observe that, by default, `.gitignore` has a list of project-specific files

- In addition, you can ignore, for example,

    - a specific folder, relative to the root directory

    - a specific file in a specific folder, relative to the root directory

    - <span style="background-color: #ffff88;">a specific file in any folder</span>

]

.pull-right[

```{r, eval = FALSE}
.Rproj.user
.Rhistory
.RData
.Ruserdata
/manuscript/
/manuscript/journals.pdf
journals.pdf               #<<
```

]

---
## Version Control &mdash; Git and Github &mdash; `.gitignore`

.pull-left[

- Observe that, by default, `.gitignore` has a list of project-specific files

- In addition, you can ignore, for example,

    - a specific folder, relative to the root directory

    - a specific file in a specific folder, relative to the root directory

    - a specific file in any folder
    
    - <span style="background-color: #ffff88;">all files with a specific extension</span>, anywhere in the project

]

.pull-right[

```{r, eval = FALSE}
.Rproj.user
.Rhistory
.RData
.Ruserdata
/manuscript/
/manuscript/journals.pdf
journals.pdf               
**.pdf
```

]

---
## Version Control &mdash; Git and Github &mdash; `.gitignore` &mdash; Notes

- There are many other pattern formats    

    - see the documentation at <https://git-scm.com/docs/gitignore>

--

- Starting to ignore a file or folder that is already being tracked requires clearing the cache

    - after changing and saving `.gitignore`, enter the following line in the `Terminal`
    - with your speficic `/path/to/file`

```yaml
git rm --cached /path/to/file

```
--

- The following command clears *all* cache

    - might be useful after changes to `.gitignore` that involves several files or folders
    - but should be used with care, on an otherwise up-to-date repository  
    
```yaml
git rm -r --cached .

```

---
class: action

## Exercises

35) Reproducibility and Version Control
- imagine that, after producing all these tables and figures, and writing up your results, you have decided to exclude journals from Oceania from analysis
   - hint: use the `filter` function in the data chunk  
   - create a new version of the manuscript
   - comit and push to GitHub
   
36) Gitignore
- stop tracking `journals.pdf`
    - change `.gitignore`
    - remove `journals.pdf` from cache
    - comit and push to GitHub


`r countdown(minutes = 5, seconds = 00, top = 0)`

---
name: part9
class: inverse, center, middle

# Part 9. Collaborating with Others

.footnote[

[Back to the contents slide](#contents-slide).

]

---

---
## Collaboration &mdash; Git and GitHub &mdash; Project Setup

- The setup depends on the users' role, on whether they are
    - the *owner* who creates the GitHub repository, or
    - the *collaborator* who is then added to that repository

--

- Once the project is setup
    - it continues to be associated with the owner's GitHub profile
    - at the same time, it is listed under the collaborator's profile as well
    - both the owner and the collaborator have the same rights, unless otherwise restricted

---
## Collaboration &mdash; Git and GitHub &mdash; Project Setup &mdash; Owner

1) The setup for the owner is largely the same as in any single-author, single-computer scenario
- following the instructions on [this slide](#rstudio-project) forward
    - to introduce version control to a local project with Git,
    - to create a remote repository for that project on GitHub, and
    - to associate the local project with the remote repository

<br>
--

2) As an additional step, the owner needs to invite their collaborator(s) to the project
- following, from the relevant GitHub repository, 

> `Settings -> Manage access -> Invite a collaborator`


---
## Collaboration &mdash; Git and GitHub &mdash; Project Setup &mdash; Collaborator


1) Notice that the remote part of the setup is done by the owner for the collaborator
- subject to acceptance of the invitation
    - invitations are available directly at <https://github.com/notifications>, but also sent via email
    - with an option to "Accept invitation"
    - on acceptance, projects appear among the repositories of the collaborator
        
--

2) The local part of the setup still needs to be done
- by creating a new RStudio project with version control
- following, from the Rstudio menu,<sup>*</sup> 

> `File -> New Project -> Version Control -> Git`

--

- the `Repository URL`, required for the above process, is the version without the `.git` extension
    - in the form of https://github.com/OWNER_USER_NAME/REPOSITORY_NAME

---
class: action

## Exercises

37) Owner Setup
- create a new version-controlled RStudio project, with Git and GitHub
- add the default R Markdown template to your project
    - hint: click `File -> New File -> R Markdown -> OK` to create the template
    - another hint: name the project and the template in a way that they are easily distinguishable from your partner's project and template

38) Invitation to Collaborate
- invite the partner in your current group as a collaborator to your new project
    - hint: you will need their username, full name, or email address to do so

39) Collaborator Setup
- accepting the invitation from your partner, do the necessary arrangements so that you can collaborate on your partner's project


`r countdown(minutes = 10, seconds = 00, top = 0)`

---
## Colloboration &mdash; Git and Github &mdash; Workflow

1) Pull

- on the Git tab in RStudio, click `Pull` to move the up-to-date records from GitHub to your computer
    - if your collaborator has not pushed anything since your last pull, you will be noticed that `Already up-to-date`.
    - collaborative projects require pulling as well as pushing because your collaborator(s) might have pushed their commits to GitHub
    - pulling frequently minimises the risk of merge conflicts

--

2) Edit and save; commit and push

- the same procedure as in any single-author, single-computer scenario
    - as described on [this slide](#git-github-basic-workflow) forward
- pushing frequently minimises the risk of merge conflicts

---
class: action

## Exercise

40) Non-simultaneous Collaboration

- take in turns with your partner to work on the same document (of the same project)
    - *owner*: edit the first header in the document (i.e., "R Markdown"), save, commit, and push
    - *owner and collaborator*: observe the changes, if any, on your own `.Rmd` file, and/or on your GitHub repository
        - click on the relevant commit message on GitHub and observe the commit
    - *collaborator*: pull, revert the header back to original, save, commit, and push

`r countdown(minutes = 5, seconds = 00, top = 0)`

--

- notice that you have not encountered any errors and/or merge conflicts
    - because everyone edited and merged with an up-to-date document
    - this is the default scenario in single-author, multiple computer scenario

---
class: action

## Exercise

41) Simultaneous Collaboration &mdash; Different Lines

- work on the same document at the same time
    - *owner*: edit the first header in the document (i.e., "R Markdown"), save, commit, and push
    - *collaborator*: edit the second header in the document (i.e., "Including Plots"), save, commit, and push
        - observe the error message that the last pusher will receive, follow the instructions on RStudio to solve the problem

`r countdown(minutes = 10, seconds = 00, top = 0)`

--

- notice that you have encountered an error but not a merge conflict
    - pulling before pushing solves the problem because the edits are not on the same line
    - the merge takes place automatically, on the *local* repository of the last pusher

---
class: action

## Exercise

42) Simultaneous Collaboration &mdash; Same Line

- work on the same document at the same time
    - *owner*: edit the first header in the document again, save, commit, and push
    - *collaborator*: edit the first header in the document as well, save, commit, and push
        - observe the error message that the last pusher will receive, follow the instructions on RStudio to solve the problem

`r countdown(minutes = 10, seconds = 00, top = 0)`

--

- notice that you have encountered not only an error but also a merge conflict
    - pulling before pushing alone does not solve the problem because the edits are on the same line
        - the conflict cannot be solved automatically &mdash; it needs human intervention
        - by pulling first, you can view the conflict directly on the file
            - marked between less than <span style="background-color: #ffff88;">&lt;</span> and greater than <span style="background-color: #ffff88;">&gt;</span> signs, divided by the equal signs
            - solution is to accept the remote version, by deleting your edit and or moving that edit to a different line
    - merging takes place on the *local* repository of the last pusher

---
## Colloboration &mdash; Git and Github &mdash; Workflow &mdash; Alternative

- The workflow above is rather simple, but has some disadvantages, including
    - not easy, albeit still possible, to see the edits of the collaborators
    - not clear who is in charge of the overall progress
    - not possible to discuss edits
    - not possible to compromise on conflicting edits

--

- An alternative workflow exits
    - work on different branches of the same project
    - version control to your own branch
    - create pull requests with comments
    - merge the branch into master


---
## Colloboration &mdash; Git and Github &mdash; Workflow &mdash; Alternative

1) Branch

- click `New Branch` on the Git tab
    - name it, and leave everything else as default
    - notice that you are now working on a new branch


2) Edit and save; commit and push

- the same procedure as in any single-author, single-computer scenario
    - as described on [this slide](#git-github-basic-workflow) forward
- notice, on GitHub, that your commit is in the new branch, while *master* remains unchanged

3) Pull request

- On GitHub, click 
> `Pull requests -> New pull request`

- choose what is to be pulled, and write a note to your collaborator who can accept or reject the merge
    - if there are merge conflicts, the collaborator solves them on GitHub before merging

---
class: action

## Exercises

43) Pull request

- create a pull request for your collaboration project
    - create a branch for yourself
    - edit any line, save, commit, and push
    - request your branch to be merged
    
44) Merging

- view the pull request of your collaborator
- take the necessary steps to merge it to *master*

`r countdown(minutes = 10, seconds = 00, top = 0)`

---
## Colloboration &mdash; Git and Github &mdash; Workflow &mdash; Notes

- It is possible to edit `.Rmd` documents directly on GitHub

    - click on any editable file, and `Edit this file`
    - commit changes, either as a direct commit or a pull request
--

- A GitHub account is enough for collaboration with co-authors who do not work with Git, R, or RStudio

    - not possible to knit to see the outcome
    - would suit co-authors whose contributions are plain text

---
class: action

## Exercises

45) GitHub edit

- create two edits on the `.Rmd` document in your collaboration project
- commit one of the edits as a direct commit
- commit the other as a pull request

`r countdown(minutes = 5, seconds = 00, top = 0)`
    
---
name: part10
class: inverse, center, middle

# Part 10. Working on a Real Project

.footnote[

[Back to the contents slide](#contents-slide).

]

---
## Real Project

- Consider converting a real project to R Markdown
    - now, in the remainder of the workshop

- Choose an existing project, preferably
    - single-authored
    - at an early stage
    - but one that you are, will be, working on

- Ask me for help
    - with no more slides to go through, I will now focus on helping you start your first project in R Markdown

---
class: inverse, center, middle

# References

.footnote[

[Back to the contents slide](#contents-slide).

]

---
## References

Allaire, J. J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins, A., Wickham, H., Cheng, J., Chang, W. and Iannone, R. (2020). [rmarkdown: Dynamic documents for R](https://cran.r-project.org/web/packages/rmarkdown/index.html). R package, version 2.1.

Blair, G., Cooper, J., Coppock, A., Humphreys, M., Rudkin, A. and Fultz, N. (2019). [fabricatr: Imagine your data before you collect it](https://cran.r-project.org/web/packages/fabricatr/index.html). R package, version 0.10.0.

Carlisle, D., Fairbairns, R., Harris, E. and Tobin, G. (2011). [setspace – Set space between lines](https://ctan.org/pkg/setspace). LaTeX package, version 6.7a.

Dowle, M. and Srinivasan, A. (2019). [data.table: Extension of 'data.frame'](https://cran.r-project.org/web/packages/data.table/index.html). R package, version 1.12.8.

Gagolewski, M. (2020). [stringi: Character String Processing Facilities](https://cran.r-project.org/web/packages/stringi/index.html). R package, version 1.4.6.

Hlavac, M. (2018). [stargazer: Well-formatted regression and summary statistics tables](https://cran.r-project.org/web/packages/stargazer/index.html). R package, version 5.2.2.

Sievert, C., Parmer, C.,  Hocking, T., Chamberlain, S., Ram, K., Corvellec, M., & Despouy, P. (2020). [plotly: Create Interactive Web Graphics via 'plotly.js'](https://cran.r-project.org/web/packages/plotly/index.html). R package, version 4.9.3.

Wickham, H. and Grolemund, G. (2019). R for data science. O'Reilly. Open access at [https://r4ds.had.co.nz](https://r4ds.had.co.nz/).

Wickham, H., Chang, W., Henry, L., Pedersen, T. L., Takahashi, K., Wilke, C., Woo, K., Yutani, H. and Dunnington, D. (2020a). [dplyr: A grammar of data manipulation](https://cran.r-project.org/web/packages/dplyr/index.html). R package, version 0.8.5.

---
## References

Wickham, H., François, W., Henry L. and Müller, K. (2020b). [ggplot2: Create elegant data visualisations using the grammar of graphics](https://cran.r-project.org/web/packages/ggplot2/index.html). R package, version 0.8.5.

Wiernik, B. M. (2020). [American Psychological Association 7th edition (no ampersand)](https://www.zotero.org/styles/apa-no-ampersand). Citation style language file, version 1.0.

Xie, Y. (2020a). [tinytex: Helper functions to install and maintain TeX Live and compile LaTeX documents](https://cran.r-project.org/web/packages/tinytex/index.html). R package, version 0.21.

Xie, Y. (2020b). [knitr: A general-purpose package for dynamic report generation in R](https://cran.r-project.org/web/packages/knitr/index.html). R package, version 1.28.

Xie, Y. (2020c). [bookdown: Authoring books and technical documents with R Markdown](https://cran.rstudio.com/web/packages/bookdown/index.html). R package, version 0.18.

Xie, Y. (2020d). [blogdown: Create blogs and websites with R Markdown](https://cran.r-project.org/web/packages/blogdown/index.html). R package, version 0.18.

Xie, Y., Allaire, J. J., and Grolemund, G. (2019). R markdown: The definitive guide. CRC Press. Open access at [https://bookdown.org/yihui/rmarkdown](https://bookdown.org/yihui/rmarkdown/).

Zhu, H. (2019). [kableExtra: Construct Complex Table with 'kable' and Pipe Syntax](https://cran.r-project.org/web/packages/kableExtra/index.html). R package, version 1.1.0.

---
class: middle, center

## The workshop ends here.
## Congratulations for making it this far, and
## thank you for joining me!

.footnote[

[Back to the contents slide](#contents-slide).

]
